---
title: "Lab 03 - Data Wrangling"
author: Jinran Yang
date: Assignment due by 11:59PM on Friday, September 14, 2018
output:
  html_document:
  toc: false
  theme: simplex
---

# Getting started  
In this exercise you will use R Studio to import and calculate summary statics of the Single Family Loan-Level Data Set, from Freddie Mac.

The Freddie Mac data includes loan-level credit performance data on fixed-rate mortgages. It includes loan-level origination and loan performance and actual loss data on Single Family mortgages acquired by Freddie Mac. The data and more details can be obtained [here](http://www.freddiemac.com/research/datasets/sf_loanlevel_dataset.html). In the User Guide, the section `File Layout & Data Dictionary` contains the description of each variable in the data sets. 

You will work with a sample of the "full" data set from 2005, 2006 and 2007. For each year, there are two files: origination and monthly performance. The data from a year contains mortgages originated in that year. The performace file contains information from the month of origination until the termination event or 09/2016, the last month in the data set.

Before you start, load the package `tidyverse` to your working space.

```{r message = FALSE}
library(tidyverse)
```

# Problem 1: Import the data
Load loan data from the folder `data_Freddie_Mac` using the `read_delim` function from the `tidyverse` package. There are two types of files--origination files and performance files-- for each year from 2005 to 2007 (six files in total).

Modify the following code chunk to load the three origination files. To help you get started, the code for the origination file for 2005 has been completed. Look at the options in `?read_delim` for an explanation of this function. Why are we trimming leading blankspaces? What are potential problems if you do not speficy the column types? 

```{r}
# Load loan origination data for 2005, 2006, and 2007
orig_2005 <-
  read_delim(
    "data_Freddie_Mac/2005/sample_orig_2005.txt.bz2",
    delim = "|",
    trim_ws = TRUE,
    col_names = c(
      "fico",
      "dt_first_pi",
      "flag_fthb",
      "dt_matr",
      "cd_msa",
      "mi_pct",
      "cnt_units",
      "occpy_sts",
      "cltv",
      "dti",
      "orig_upb",
      "ltv",
      "int_rt",
      "channel",
      "ppmt_pnlty",
      "prod_type",
      "st",
      "prop_type",
      "zipcode",
      "id_loan",
      "loan_purpose",
      "orig_loan_term",
      "cnt_borr",
      "seller_name",
      "servicer_name",
      "flag_sc"
    ),
    col_types = cols(
      "i",
      "i",
      "c",
      "i",
      "c",
      "d",
      "i",
      "c",
      "d",
      "i",
      "i",
      "i",
      "d",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "i",
      "i",
      "c",
      "c",
      "c"
    )
  )
orig_2006 <- read_delim("data_Freddie_Mac/2006/sample_orig_2006.txt.bz2", delim = "|",
    trim_ws = TRUE,
    col_names = c(
      "fico",
      "dt_first_pi",
      "flag_fthb",
      "dt_matr",
      "cd_msa",
      "mi_pct",
      "cnt_units",
      "occpy_sts",
      "cltv",
      "dti",
      "orig_upb",
      "ltv",
      "int_rt",
      "channel",
      "ppmt_pnlty",
      "prod_type",
      "st",
      "prop_type",
      "zipcode",
      "id_loan",
      "loan_purpose",
      "orig_loan_term",
      "cnt_borr",
      "seller_name",
      "servicer_name",
      "flag_sc"
    ),
    col_types = cols(
      "i",
      "i",
      "c",
      "i",
      "c",
      "d",
      "i",
      "c",
      "d",
      "i",
      "i",
      "i",
      "d",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "i",
      "i",
      "c",
      "c",
      "c"
    )
  )
orig_2007 <- read_delim("data_Freddie_Mac/2007/sample_orig_2007.txt.bz2",  delim = "|",
    trim_ws = TRUE,
    col_names = c(
      "fico",
      "dt_first_pi",
      "flag_fthb",
      "dt_matr",
      "cd_msa",
      "mi_pct",
      "cnt_units",
      "occpy_sts",
      "cltv",
      "dti",
      "orig_upb",
      "ltv",
      "int_rt",
      "channel",
      "ppmt_pnlty",
      "prod_type",
      "st",
      "prop_type",
      "zipcode",
      "id_loan",
      "loan_purpose",
      "orig_loan_term",
      "cnt_borr",
      "seller_name",
      "servicer_name",
      "flag_sc"
    ),
    col_types = cols(
      "i",
      "i",
      "c",
      "i",
      "c",
      "d",
      "i",
      "c",
      "d",
      "i",
      "i",
      "i",
      "d",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "c",
      "i",
      "i",
      "c",
      "c",
      "c"
    )
  )
```

Modify the code chunk below to load the three performance data files. Follow the User Guide to determine the variable types. Note that only the first 22 variables listed in the User Guide are included in the data file. 

```{r}
# Load loan performance data for 2005, 2006, and 2007
svcg_2005  <-
  read_delim(
    "data_Freddie_Mac/2005/sample_svcg_2005.txt.bz2",
    delim = "|",
    trim_ws = TRUE,
    col_names = c(
      "id_loan", 
      "svcg_cycle", 
      "current_upb", 
      "delq_sts", 
      "loan_age", 
      "mths_remng", 
      "repch_flag", 
      "flag_mod",
      "cd_zero_bal", 
      "dt_zero_bal", 
      "current_int_rt", 
      "non_int_brng_upb", 
      "dt_lst_pi", 
      "mi_recoveries",
      "net_sale_proceeds", 
      "non_mi_recoveries", 
      "expenses", 
      "legal_costs", 
      "maint_pres_costs", 
      "taxes_ins_costs",
      "misc_costs", 
      "actual_loss"
    ),
    col_types = cols( 
      "c",
      "i",
      "d",
      "c",
      "d",
      "i",
      "c",
      "c",
      "i",
      "i",
      "d",
      "d",
      "i",
      "d",
      "c",
      "d",
      "d",
      "d",
      "d",
      "d",
      "d",
      "d"
    )
  )
  

svcg_2006 <- read_delim("data_Freddie_Mac/2006/sample_svcg_2006.txt.bz2",  delim = "|",
    trim_ws = TRUE,
    col_names = c(
      "id_loan", 
      "svcg_cycle", 
      "current_upb", 
      "delq_sts", 
      "loan_age", 
      "mths_remng", 
      "repch_flag", 
      "flag_mod",
      "cd_zero_bal", 
      "dt_zero_bal", 
      "current_int_rt", 
      "non_int_brng_upb", 
      "dt_lst_pi", 
      "mi_recoveries",
      "net_sale_proceeds", 
      "non_mi_recoveries", 
      "expenses", 
      "legal_costs", 
      "maint_pres_costs", 
      "taxes_ins_costs",
      "misc_costs", 
      "actual_loss"
    ),
    col_types = cols( 
      "c",
      "i",
      "d",
      "c",
      "d",
      "i",
      "c",
      "c",
      "i",
      "i",
      "d",
      "d",
      "i",
      "d",
      "c",
      "d",
      "d",
      "d",
      "d",
      "d",
      "d",
      "d"
    )
  )
svcg_2007 <- read_delim("data_Freddie_Mac/2007/sample_svcg_2007.txt.bz2",   delim = "|",
    trim_ws = TRUE,
    col_names = c(
      "id_loan", 
      "svcg_cycle", 
      "current_upb", 
      "delq_sts", 
      "loan_age", 
      "mths_remng", 
      "repch_flag", 
      "flag_mod",
      "cd_zero_bal", 
      "dt_zero_bal", 
      "current_int_rt", 
      "non_int_brng_upb", 
      "dt_lst_pi", 
      "mi_recoveries",
      "net_sale_proceeds", 
      "non_mi_recoveries", 
      "expenses", 
      "legal_costs", 
      "maint_pres_costs", 
      "taxes_ins_costs",
      "misc_costs", 
      "actual_loss"
    ),
    col_types = cols( 
      "c",
      "i",
      "d",
      "c",
      "d",
      "i",
      "c",
      "c",
      "i",
      "i",
      "d",
      "d",
      "i",
      "d",
      "c",
      "d",
      "d",
      "d",
      "d",
      "d",
      "d",
      "d"
    )
  )
```


# Problem 2: Append the data frames 
Using the `bind_rows` command from the `dplyr` package, combine the 2005-2007 origination data into a single data frame, called `orig`. Limit this data frame to only include the following variables: `id_loan, orig_upb, fico, cltv, ltv, dti, int_rt, st`. Use the pipe operator `%>%` to perform the sequence of `dplyr` commands to create this data frame.

Similarly, combine the 2005-2007 performance data into a single data frame, called `svcg`. Limit this data frame to only include the following variables: `id_loan, svcg_cycle, cd_zero_bal, delq_sts, flag_mod, actual_loss`.

Finally, erase from the R working environment the six year-specific data frames used to create the `orig` and `svcg` data frames.

```{r}
# Combined origination and performance data
orig <-bind_rows(orig_2007,orig_2006,orig_2005)%>%
select(id_loan, orig_upb, fico, cltv, ltv, dti, int_rt, st)

svcg <- bind_rows(svcg_2007,svcg_2006,svcg_2005)%>%
select(id_loan, svcg_cycle, cd_zero_bal, delq_sts, flag_mod, actual_loss)

# Delete component files
rm(svcg_2007,svcg_2006,svcg_2005,orig_2007,orig_2006,orig_2005)
```


# Problem 3: Join the datasets
Our final goal is to analyze the performace of loans in relation to some of its origination characteristics, such as credit score. For this, the next step is to join the origination and performance datasets using the `left_join()` command from the `dplyr` package. Name the resulting joined data frame `orig_svcg`.

```{r}
orig_svcg <- left_join(svcg,orig,by="id_loan")
```


# Problem 4: Flag of duplicated loan IDs
Take a look at the `orig_svcg` data frame. It has multiple entries for each mortgage. This is useful for summarizing performance statistics, but will overcount statistics by loan. 

Add a new variable to `orig_svcg` that flags duplicate records. Name this new variable `duplicated`. The variable should take a value of `FALSE` for the first month in which a loan is observed. The variable should take a value of `TRUE` for all subsequent months in which the loan is observed.

**Hint:** You may want to use the `dplyr` functions `arrange`, `mutate`, and `duplicated`.

```{r}
# Arrange observations in order of svcg_cyle
# The first occurence of id_loan will have a duplicated value of FALSE, and all others will have values of TRUE
orig_svcg <- orig_svcg %>%
  arrange(svcg_cycle)%>%
  mutate(duplicated = duplicated(id_loan,incomparables = FALSE))

```


# Problem 5: Flag defaulted loans
In the rest of the analysis, consider loans that are more than 60 days late or those with a delinquency status of "R" (REO Acquisition) as in "default." In this case, we will say that the loan has reached a delinquency status of "D60." Make sure you understand how the delinquency status variable `delq_sts` is defined in order to get the desired measure of default.

Create a variable that equals 1 if a loan has ever defaulted. Follow these steps:

1. Create a variable named `default` that identifies whether a loan is in default (D60) in any month. Use the command `mutate()` that you learned in class. 
2. Crete a variable by loan named `default_byloan` that equals 1 if the loan was defaulted at any point. Use the commands `mutate()` and `group_by`.
3. After using `group_by`, you need to ungroup the data frame. To do this, run the code `orig_svcg <- ungroup(orig_svcg)`.

```{r}
orig_svcg<-orig_svcg%>%
  mutate(default = ifelse(delq_sts>=2|delq_sts == "R",1,0))%>%
  group_by(id_loan)%>%
  mutate(default_byloan = max(default,na.rm = T))%>%
  ungroup(orig_svcg)



```


# Problem 6: Summary statistics
Build a table of summary statistics for the Freddie Mac data. 

Calculate the following summary statistics, and use the duplicated flag as needed. Print the statistics into a table called `Summary Statistics - Freddie Mac`.

1. Loan count 
2. Average original UPB (Unpaid principal balance) (in US$) 
3. Credit score, as a weighted average using the original UPB
4. Original combined loan-to-value ratio as a weighted average using the original UPB (CLTV: the ratio of all loans secured by a property to the property's value)
5. Original debt-to-income ratio, as a weighted average using the original UPB (DTI: percentage of a consumer's monthly gross income that goes toward paying debts)
6. Ever D60 % (Count of loans that reached a delinquency status of D60, including at REO Acquisition, divided by Loan Count - use the `default_byloan` variable)

```{r}
num_load<- orig_svcg %>%         ###for question 6
  select(default_byloan,id_loan)%>%
  filter(default_byloan == 1)

orig_svcg6<- orig_svcg%>%
  filter(duplicated == 0)
  
table1<-summarise(orig_svcg6,Table_name = "Summary Statistics - Freddie Mac",
          loan_count= length(unique(orig_svcg$id_loan)),
          Ave_Ori_UPB = mean(orig_svcg6$orig_upb),
          Credit_Score = weighted.mean(orig_svcg6$fico,orig_svcg6$orig_upb,na.rm=T),
          Ratio = weighted.mean(orig_svcg6$cltv,orig_svcg6$orig_upb,na.rm=T),
          Ori_debt_to_income_ratio = weighted.mean(orig_svcg6$dti,orig_svcg6$orig_upb,na.rm=T),
          Ever_D60 = length(unique(num_load$id_loan))/length(unique(orig_svcg$id_loan)))
table1
```
# Problem 7: Save the data frame
Using the R built-in command `save()`, save the data frame `orig_svcg` to a file named `cleaned_Freddie_Mac.Rdata` in the `lab-03` folder.

```{r}
save(orig_svcg,file = "cleaned_Freddie_Mac.Rdata")
```

